---
title: "Metabolomics_human_II"
output: html_document
date: "2024-12-03"
editor_options: 
  chunk_output_type: inline
---



# Library 
```{r}
library(omu)
library(tidyverse)

count_df <- read_metabo(filepath = "/Users/abnerportillo/Downloads/metabolomics_human.csv") #Started w/ 9671 observations

meta_data_df <- read.csv("/Users/abnerportillo/Downloads/meta_data_human.csv")


count_df <- assign_hierarchy(count_data = count_df, keep_unknowns = TRUE, identifier = "KEGG")

sum(is.na(count_df$Class))

count_df[is.na(count_df)] <- "Other"





```

#Data for PCA plot (dont run unless want to see PCA plot)
```{r}

count_df_log <- log(count_df[,3:79]) # transforming data 

count_df_log <- cbind(count_df[,1:2], count_df_log) #combining data 




#want to add class columns to count_df_log for later analysis
count_df_log <- assign_hierarchy(count_data = count_df_log, keep_unknowns = TRUE, identifier = "KEGG")


# Keep rows where all values in columns 3 to 79 are not zero
count_df_log <- count_df_log[apply(count_df_log[, 3:79], 1, function(row) all(row != 0)), ]




```

#PCA plot
```{r}
PCA <- PCA_plot(count_data = count_df_log, metadata = meta_data_df, variable ="Arm",color = "Arm",response_variable = "Metabolite") +theme_bw() + theme(panel.grid = element_blank())

PCA
```

# Important for the comparisons (In meta-data)
```{r}
#making new columns for comparisons

meta_data_df$Condition<-factor(paste(meta_data_df$Arm, meta_data_df$Time))

```

# NIF baseline vs Week 5
```{r}
NIF_b_w5 <- omu_summary(

  count_data = count_df,
  metadata = meta_data_df,
  numerator = "Non-Intermittent B0",
  denominator = "Non-Intermittent D5",
  response_variable = "Metabolite",
  Factor = "Condition",
  log_transform = TRUE,
  p_adjust = "BH",
  test_type = "mwu"
)

### See what metabolites are significant 

NIF_b_w5$diffexpressed <- "NO"
# if Foldchange >1 and pvalue < 0.05, set as "UP"
NIF_b_w5$diffexpressed[NIF_b_w5$Fold_Change > 1.0 & NIF_b_w5$pval < 0.05] <- "UP"
# if Foldchange < -1 and pvalue < 0.05, set as "DOWN"
NIF_b_w5$diffexpressed[NIF_b_w5$Fold_Change < 1.0 & NIF_b_w5$pval < 0.05] <- "DOWN"

#How many significant metabolites? 
NIF_b_w5 %>% count(diffexpressed)

summary(NIF_b_w5$Class)## Finding what other classes to subset by

#What are the metabolites that are significant by Class 

 NIF_b_w5_class<- NIF_b_w5[c("Metabolite", "KEGG","Class", "diffexpressed", "pval","Fold_Change")]
 NIF_b_w5_class <- NIF_b_w5_class[which(NIF_b_w5_class[,"pval"] <= 0.05),]
 
 summary(NIF_b_w5_class)
 
```


```{r}
#Subsetting for lipids
NIF_b_w5_lipids<- subset(NIF_b_w5, Class=="Lipids")
NIF_b_w5_lipids <- NIF_b_w5_lipids[which(NIF_b_w5_lipids[,"pval"] <= 0.05),]

NIF_b_w5_lipids %>% count(diffexpressed)


# Phytochemical compounds, Carbohydrates, Nucleic acids, Peptides, Organic acids, Hormones and Transmitters, Antibiotics, Vitamins and Cofactors, Bioactive peptides

# Subsetting for Carbohydrates 
NIF_b_w5_carbs <- subset(NIF_b_w5, Class == "Carbohydrates")
NIF_b_w5_carbs <- NIF_b_w5_carbs[which(NIF_b_w5_carbs[,"pval"] <= 0.05),]
# None where significant

#Subsetting for Phytochemical compounds
NIF_b_w5_phy_chem <- subset(NIF_b_w5, Class == "Phytochemical compounds")
NIF_b_w5_phy_chem <- NIF_b_w5_phy_chem[which(NIF_b_w5_phy_chem[,"pval"] <= 0.05),]
NIF_b_w5_phy_chem  %>% count(diffexpressed)

NIF_b_w5_n_acid <- subset(NIF_b_w5, Class == "Nucleic acids")
NIF_b_w5_n_acid <-NIF_b_w5_n_acid[which(NIF_b_w5_n_acid[,"pval"] <= 0.05), ]

NIF_b_w5_peptides <- subset(NIF_b_w5, Class == "Peptides")
NIF_b_w5_peptides <- NIF_b_w5_peptides[which(NIF_b_w5_peptides[, "pval"] <= 0.05),]



```

# NIF baseline vs NIF 3 month
```{r}
NIF_b_3m <- omu_summary(

  count_data = count_df,
  metadata = meta_data_df,
  numerator = "Non-Intermittent B0",
  denominator = "Non-Intermittent F3",
  response_variable = "Metabolite",
  Factor = "Condition",
  log_transform = TRUE,
  p_adjust = "BH",
  test_type = "mwu"
  
)

### See what metabolites are significant 

NIF_b_3m$diffexpressed <- "NO"
# if Foldchange >1 and pvalue < 0.05, set as "UP"
NIF_b_3m$diffexpressed[NIF_b_3m$Fold_Change > 1.0 & NIF_b_3m$pval < 0.05] <- "UP"
# if Foldchange < -1 and pvalue < 0.05, set as "DOWN"
NIF_b_3m$diffexpressed[NIF_b_3m$Fold_Change < 1.0 & NIF_b_3m$pval < 0.05] <- "DOWN"

#How many significant metabolites? 
NIF_b_3m %>% count(diffexpressed)

summary(NIF_b_3m$Class)## Finding what other classes to subset by

#What are the metabolites that are significant by Class 

 NIF_b_3m_class<- NIF_b_3m[c("Metabolite", "KEGG","Class", "diffexpressed", "pval","Fold_Change")]
 NIF_b_3m_class <- NIF_b_3m_class[which(NIF_b_3m_class[,"pval"] <= 0.05),]
 
 
### see what class is affected
 summary(NIF_b_3m_class)

```
# IF baseline vs Week 5
```{r}
IF_b_w5 <- omu_summary(

  count_data = count_df,
  metadata = meta_data_df,
  numerator = "Intermittent Fasting B0",
  denominator = "Intermittent Fasting D5",
  response_variable = "Metabolite",
  Factor = "Condition",
  log_transform = TRUE,
  p_adjust = "BH",
  test_type = "mwu"
  
)

### See what metabolites are significant 

IF_b_w5$diffexpressed <- "NO"
# if Foldchange >1 and pvalue < 0.05, set as "UP"
IF_b_w5$diffexpressed[IF_b_w5$Fold_Change > 1.0 & IF_b_w5$pval < 0.05] <- "UP"
# if Foldchange < -1 and pvalue < 0.05, set as "DOWN"
IF_b_w5$diffexpressed[IF_b_w5$Fold_Change < 1.0 & IF_b_w5$pval < 0.05] <- "DOWN"

#How many significant metabolites? 
IF_b_w5 %>% count(diffexpressed)

summary(IF_b_w5$Class)## Finding what other classes to subset by

#What are the metabolites that are significant by Class 

IF_b_w5_class<-  IF_b_w5[c("Metabolite", "KEGG","Class", "diffexpressed", "pval","Fold_Change")]
 IF_b_w5_class <- IF_b_w5_class[which(IF_b_w5_class[,"pval"] <= 0.05),]
 
### see what class is affected
 summary(IF_b_w5_class)

 


```
# IF baseline vs month 3

```{r}
IF_b_3m <- omu_summary(

  count_data = count_df,
  metadata = meta_data_df,
  numerator = "Intermittent Fasting B0",
  denominator = "Intermittent Fasting F3",
  response_variable = "Metabolite",
  Factor = "Condition",
  log_transform = TRUE,
  p_adjust = "BH",
  test_type = "mwu"
  
)

### See what metabolites are significant 

IF_b_3m$diffexpressed <- "NO"
# if Foldchange >1 and pvalue < 0.05, set as "UP"
IF_b_3m$diffexpressed[IF_b_3m$Fold_Change > 1.0 & IF_b_3m$pval < 0.05] <- "UP"
# if Foldchange < -1 and pvalue < 0.05, set as "DOWN"
IF_b_3m$diffexpressed[IF_b_3m$Fold_Change < 1.0 & IF_b_3m$pval < 0.05] <- "DOWN"

#How many significant metabolites? 
IF_b_3m %>% count(diffexpressed)

summary(IF_b_3m$Class)## Finding what other classes to subset by

#What are the metabolites that are significant by Class 

IF_b_3m_class<-  IF_b_3m[c("Metabolite", "KEGG","Class", "diffexpressed", "pval", "Fold_Change")]
 IF_b_3m_class <- IF_b_3m_class[which(IF_b_3m_class[,"pval"] <= 0.05),]
 
### see what class is affected
 summary(IF_b_3m_class)

```
identify common and unique metabolites

#Upregulated Venn diagram for NIF
```{r}
#Subset the metabolites that are UP regulated into own df 

### Use the class df 
Upregulated_NIF_w5 <- subset(NIF_b_w5_class, diffexpressed == "UP")

Upregulated_NIF_w5$Time_point <- "Week 5"


Upregulated_NIF_m3 <- subset(NIF_b_3m_class, diffexpressed == "UP")

Upregulated_NIF_m3$Time_point <- "Month 3"




library(VennDiagram)
venn.diagram(
  x = list(Upregulated_NIF_w5$Metabolite, Upregulated_NIF_m3$Metabolite),
  category.names = c("Week 5", "Month 3"),
  filename = "venn_metabolites_NIF_all_U.png"
)

common_metabolites_NIF_all_U <- Reduce(intersect,list(Upregulated_NIF_w5$Metabolite, Upregulated_NIF_m3$Metabolite))

print(common_metabolites_NIF_all_U)

```
Downregulated Venn diagram for NIF 
```{r}
Downregulated_NIF_w5 <- subset(NIF_b_w5, diffexpressed == "DOWN")

Downregulated_NIF_w5$Time_point <- "Week 5"

Downregulated_NIF_m3 <- subset(NIF_b_3m, diffexpressed == "DOWN")

Downregulated_NIF_m3$Time_point <- "Month 3"

library(VennDiagram)
venn.diagram(
  x = list(Downregulated_NIF_w5$Metabolite, Downregulated_NIF_m3$Metabolite),
  category.names = c("Week 5", "Month 3"),
  filename = "venn_metabolites_NIF_all_D.png"
)

common_metabolites_NIF_all_D <- Reduce(intersect,list(Downregulated_NIF_w5$Metabolite, Downregulated_NIF_m3$Metabolite ))

print(common_metabolites_NIF_all_D)



```

# significant metabolites seen in week 5 and month 3 

"Adenosine 5'-Diphosphate" "Carnosol" "Niacinamide"  

"Boldine"  "PE-NMe2(16:0/18:1(11Z))"

Upregulated metabolites venn diagram IF
```{r}
Upregulated_IF_w5 <- subset(IF_b_w5_class, diffexpressed == "UP")

Upregulated_IF_m3 <- subset(IF_b_3m_class, diffexpressed =="UP")

library(VennDiagram)
venn.diagram(
  x = list(Upregulated_IF_w5$Metabolite, Upregulated_IF_m3$Metabolite),
  category.names = c("Week 5", "Month 3"),
  filename = "venn_metabolites_IF_U.png"
)

common_metabolites_IF_all_U <- Reduce(intersect,list(Upregulated_IF_w5$Metabolite, Upregulated_IF_m3$Metabolite ))

print(common_metabolites_IF_all_U)

```
#Downregulated metabolites venn Diagram IF
```{r}
Downregulated_IF_w5 <- subset(IF_b_w5_class, diffexpressed == "DOWN")

Downregulated_IF_w5$Time_point <- "Week 5"

Downregulated_IF_m3 <- subset(IF_b_3m_class, diffexpressed == "DOWN")

Downregulated_IF_m3$Time_point <- "Month 3"

library(VennDiagram)
venn.diagram(
  x = list(Downregulated_IF_w5$Metabolite, Downregulated_IF_m3$Metabolite),
  category.names = c("Week 5", "Month 3"),
  filename = "venn_metabolites_IF_all_D.png"
)

common_metabolites_IF_all_D <- Reduce(intersect,list(Downregulated_IF_w5$Metabolite, Downregulated_IF_m3$Metabolite ))

print(common_metabolites_IF_all_D)
```

##Combining only significant metabolites that are seen in both week 5 and month 3

"1-Octadecanoyl-2-eicosanoyl-glycero-3-phosphoserine"
"Adenosine monophosphate"                            
 "Raffinose"         
"17alpha-Hydroxypregnenolone"              
"N-(octadecanoyl)-4R-hydroxysphinganine"   
 "PC(20:1(11Z)/22:6(4Z,7Z,10Z,13Z,16Z,19Z))"
 "PE-NMe(15:0/22:2(13Z,16Z))"               
"TG(14:0/14:0/20:5(5Z,8Z,11Z,14Z,17Z))"  


#Track changes across time points 
#Boxplots
```{r}
# Changes across Condition (Intermittent fasting vs Non-Intermmittent fasting)

boxplot_metabolites_condition<- plot_boxplot(count_data = count_df, metadata = meta_data_df, log_transform = TRUE, Factor = "Arm", response_variable = "Metabolite", fill_list = c("darkgoldenrod1", "dodgerblue2"), aggregate_by = "Class")

print(boxplot_metabolites_condition)

```

```{r}
# Change across time 
boxplot_metabolites_time <- plot_boxplot(count_data = count_df, metadata = meta_data_df, aggregate_by = "Class", Factor = "Time",log_transform = TRUE, fill_list =c("darkgoldenrod1", "dodgerblue2","red","darkorange","green","purple","pink","black") )

print(boxplot_metabolites_time)
```
### I know what specific class of metabolites change just have to extract only significant values. 
sig_threshold is taking into account padj not pval, need to get it to use pval so I only focus on those that are differentially expressed. This is looking at all of them. 
#Count fold function

```{r Count fold change}
### Function count_fold changes ( changing the requirments from padj to pval and Log2FoldChange to Fold_Change)

count_fold_changes <- function(count_data, column, sig_threshold, keep_unknowns){

  if(is.null(count_data$pval)==TRUE){

    stop("count_data must be the output of omu_summary or omu_anova and have a padj column
    and a log2FoldChange column")

  }

  if(any(names(count_data) %in% column)==FALSE){

    stop("count_data is missing metabolite metadata. Did you forget to use assign_hierarchy?")

  }


class(count_data) <- "data.frame"

   Fold_Change <- neg <- NULL

  count_data <- count_data[which(count_data[,"pval"] <= sig_threshold),]

  count_data <- count_data %>% group_by(count_data[,column]) %>%
   mutate(Significant_Changes = sum(Fold_Change > 1),
          neg = sum(Fold_Change<1))

 count_data <- count_data[,c(column, "Significant_Changes", "neg")]
 count_data$neg <- count_data$neg * -1
 output <- count_data[,c(1,3)]
 count_data <- count_data[,c(1,2)]
 colnames(output)[2] <- "Significant_Changes"
 count_data <- rbind(count_data, output)

 count_data$color <- ifelse(count_data$Significant_Changes < 0, "Decrease","Increase")

 count_data <- count_data[apply(count_data[2],1,function(z) !any(z==0)),]


  if(keep_unknowns==TRUE){
    count_data = count_data
    }else if(keep_unknowns==FALSE){
     count_data <- drop_na(count_data)
    }
  unique(count_data[])

}

```
# Individual classes 
```{r}
##Week 5
fold_change_counts_w5 <- count_fold_changes(count_data = NIF_b_w5,
column = "Class", sig_threshold = 0.05, keep_unknowns = "FALSE")

library(ggplot2)
Class_Bar_Plot_w5 <- plot_bar(fc_data = fold_change_counts_w5, fill = c("dodgerblue2", "firebrick2"), outline_color = c("black", "black"), size = c(1,1)) + labs(x = "Class") +theme(panel.grid =element_blank(), axis.text.x = element_text( size = 10 )
)
print(Class_Bar_Plot_w5)



#### Plot as percentage 
#### Understand wh
```

```{r}
fold_change_counts_3m <- count_fold_changes(count_data = NIF_b_3m,
column = "Class", sig_threshold = 0.05, keep_unknowns = "FALSE")

library(ggplot2)
Class_Bar_Plot_3m <- plot_bar(fc_data = fold_change_counts_3m, fill = c("dodgerblue2", "firebrick2"), outline_color = c("black", "black"), size = c(1,1)) + labs(x = "Class") +theme(panel.grid =element_blank(), axis.text.x = element_text( size = 10 )
)
print(Class_Bar_Plot_3m)
```

```{r}
##Week 5
fold_change_counts_w5_IF <- count_fold_changes(count_data = IF_b_w5,
column = "Class", sig_threshold = 0.05, keep_unknowns = "FALSE")

library(ggplot2)
Class_Bar_Plot_w5_IF <- plot_bar(fc_data = fold_change_counts_w5_IF, fill = c("dodgerblue2", "firebrick2"), outline_color = c("black", "black"), size = c(1,1)) + labs(x = "Class") +theme(panel.grid =element_blank(), axis.text.x = element_text( size = 10)
)
print(Class_Bar_Plot_w5_IF)
```

```{r}

fold_change_counts_3m_IF <- count_fold_changes(count_data = IF_b_3m,
column = "Class", sig_threshold = 0.05, keep_unknowns = "FALSE")

library(ggplot2)
Class_Bar_Plot_3m_IF <- plot_bar(fc_data = fold_change_counts_3m_IF, fill = c("dodgerblue2", "firebrick2"), outline_color = c("black", "black"), size = c(1,1)) + labs(x = "Class") +theme(panel.grid =element_blank(), axis.text.x = element_text( size = 10 )
)
print(Class_Bar_Plot_3m_IF)

```
```{r}
#Other ways to visualize the data
#Statistcal tests for the boxplots
# What do I do next??


```

#Scatterplot of each patient showing trend across time
will plot only the significant metabolites from baseline vs week 5, and baseline month 3 for now. 
```{r}

NIF_b_3m_class$Time_point <- "Month 3"
NIF_b_3m_class$Condition <- "Non-Intermittent Fasting"

NIF_b_w5_class$Time_point <- "Week 5"
NIF_b_w5_class$Condition <- "Non-Intermittent Fasting"

IF_b_3m_class$Time_point <-"Month 3"
IF_b_3m_class$Condition <- "Intermittent Fasting"

IF_b_w5_class$Time_point <- "Week 5"
IF_b_w5_class$Condition <- "Intermmittent Fasting"



#Combine dataframes first

#NIF significant metabolites both UP and DOWN regulated
NIF_combined <- rbind(NIF_b_w5_class, NIF_b_3m_class)

#IF significant metabolites both UP and DOWN regulated
IF_combined <- rbind(IF_b_w5_class, IF_b_3m_class)


```

```{r}

### Add a replicate column to metadata to differntiate bewtwen them 

meta_data_df <- meta_data_df %>%
  mutate(
    Replicate = ifelse(str_detect(Sample_ID, "_2$"), "Replicate", "Original")
  )

  ### So I will make new patient column to join for the plot 
meta_data_df <-  meta_data_df %>% mutate(Patient = str_split_fixed(Sample_ID, "_", 2)[,1])
  

  
```

```{r}
only_count <- read_csv(file = "/Users/abnerportillo/Downloads/only_count_data_human.csv") 
## going to have to subset for only those that are significant metabolites 
rownames(only_count) <- only_count$Metabolite


#Filter for significant metabolites in NIF
tmp_NIF <- only_count[NIF_combined$Metabolite,]

long_count_NIF <- pivot_longer(tmp_NIF, cols = starts_with("S"), 
                           names_to = "Sample",
                           values_to = "Count")

## Dont join but instead filter for significant metabolites 

#Filter for significant metabolites in IF 
tmp_IF <- only_count[IF_combined$Metabolite,]

long_count_IF <- pivot_longer(tmp_IF, cols = starts_with("S"),
                           names_to = "Sample",
                           values_to = "Count")


```

#Merge long df's with meta data 
```{r}

### I have to merge only with those that are not intermittent fasting in metadata
merged_data_NIF <- long_count_NIF %>%
  left_join(meta_data_df, by = "Sample")

#Filter out for only NIF arm 
merged_data_NIF <- merged_data_NIF %>% filter(Arm == "Non-Intermittent")

print(merged_data_NIF)

### Have to merge with only those that are IF in metadata
merged_data_IF <- long_count_IF %>%
  left_join(meta_data_df, by = "Sample")

#Filter out for only IF arm 
merged_data_IF <- merged_data_IF %>% filter(Arm == "Intermittent Fasting")

print(merged_data_IF) 

```

```{r}
#Filter out for only significant metaboltes seen in both week 5 and month 3

selected_metabolites_NIF <- c("Adenosine 5'-Diphosphate", "Carnosol", "Niacinamide","Boldine",  "PE-NMe2(16:0/18:1(11Z))"  )

# Filter the data frame
merged_data_NIF_only_sig <- merged_data_NIF %>%
 filter(Metabolite %in% selected_metabolites_NIF)



selelcted_metabolites_IF <- c("1-Octadecanoyl-2-eicosanoyl-glycero-3-phosphoserine",
"Adenosine monophosphate", "Raffinose" ,"17alpha-Hydroxypregnenolone" ,       "N-(octadecanoyl)-4R-hydroxysphinganine", "PC(20:1(11Z)/22:6(4Z,7Z,10Z,13Z,16Z,19Z))",
 "PE-NMe(15:0/22:2(13Z,16Z))" , "TG(14:0/14:0/20:5(5Z,8Z,11Z,14Z,17Z))" )

merged_data_IF_only_sig <- merged_data_IF %>%
  filter(Metabolite %in% selelcted_metabolites_IF)



# Filter based on condition 



```


#Plotting 
```{r  plot wider, fig.width= 15, fig.height= 7}
# NIF patients 

# To fix issue with graphs not showing up correctly 



ggplot(merged_data_NIF_only_sig, 
       aes(x = Time, y = Count, color = interaction(Patient, Replicate), group = interaction(Patient, Metabolite, Replicate))) +
  geom_line() +
  geom_point() +
  facet_wrap(~Metabolite, scales = "free_y") +
  theme_minimal() +
  labs(
    title = "Metabolite Counts Over Time",
    x = "Time",
    y = "Count",
    color = "Patient (Replicate)"  # Legend title
  ) +
  scale_x_discrete(labels = c("B0", "D1", "D1A", "D1B", "D5", "D7", "F3", "F6", "F9", "F12")) 




```

```{r plot wider, fig.width= 13, fig.height=7}
# For IF patients
ggplot(merged_data_IF_only_sig,  aes(x = Time, y = Count, color = interaction(Patient, Replicate), group = interaction(Patient, Metabolite, Replicate))) +
  geom_line() +
  geom_point() +
  facet_wrap(~Metabolite, scales = "free_y") +
  theme_minimal() +
  labs(
    title = "Metabolite Counts Over Time",
    x = "Time",
    y = "Count",
    color = "Patient (Replicate)"  # Legend title
  ) +
  scale_x_discrete(labels = c("B0", "D1", "D1A", "D1B", "D5", "D7", "F3", "F6", "F9", "F12"))

```
# Have to fix all of the x axis labels on each figure so i can see all of the time oints and 



#Week 5 signifcant changes in percentages 
```{r}
# significant changes as percentages table 
NIF_w5_percent_changes <- ra_table(fc_data = fold_change_counts_w5, variable = "Class")

# Pie chart with percentages of compounds that increased significantly (NIF)
pie_chart_NIF_w5_inc <- pie_chart(ratio_data = NIF_w5_percent_changes, variable = "Class", column = "Increase", color = "black") +
  labs(title = "Percentage of compounds that increased per class level NIF") +
  guides(fill = guide_legend(title = "Compound classes"))  # Change legend title
print(pie_chart_NIF_w5_inc)

# Pie chart with percentages of compounds that decreased significantly (NIF)
pie_chart_NIF_w5_dec <- pie_chart(ratio_data = NIF_w5_percent_changes, variable = "Class", column = "Decrease", color = "black") +
  labs(title = "Percentage of compounds that decreased per class level NIF") +
  guides(fill = guide_legend(title = "Compound classes"))
print(pie_chart_NIF_w5_dec)

# Pie chart with all significant changes (NIF)
pie_chart_NIF_w5_all <- pie_chart(ratio_data = NIF_w5_percent_changes, variable = "Class", column = "Significant_Changes", color = "black") +
  labs(title = "Percentage of compounds that are significant per class level NIF") +
  guides(fill = guide_legend(title = "Compound classes"))
print(pie_chart_NIF_w5_all)

#Intermittent fasting w5 changes as percentages
IF_w5_percent_changes <- ra_table(fc_data = fold_change_counts_w5_IF, variable = "Class")


# Pie chart with percentages of compounds that increased significantly (IF)
pie_chart_IF_w5_inc <- pie_chart(ratio_data = IF_w5_percent_changes, variable = "Class", column = "Increase", color = "black") +
  labs(title = "Percentage of compounds that increased per class level IF") +
  guides(fill = guide_legend(title = "Compound classes"))
print(pie_chart_IF_w5_inc)

# Pie chart with percentages of compounds that decreased significantly (IF)
pie_chart_IF_w5_dec <- pie_chart(ratio_data = IF_w5_percent_changes, variable = "Class", column = "Decrease", color = "black") +
  labs(title = "Percentage of compounds that decreased per class level IF") +
  guides(fill = guide_legend(title = "Compound classes"))
print(pie_chart_IF_w5_dec)

# Pie chart with all significant changes (IF)
pie_chart_IF_w5_all <- pie_chart(ratio_data = IF_w5_percent_changes, variable = "Class", column = "Significant_Changes", color = "black") +
  labs(title = "Percentage of compounds that are significant per class level IF") +
  guides(fill = guide_legend(title = "Compound classes"))
print(pie_chart_IF_w5_all)

```


# Month 3 significant changes in percentages 
```{r}

##NIF percent change table 
NIF_m3_percent_changes <- ra_table(fc_data = fold_change_counts_3m, variable = "Class")

# pie chart with percentages of compounds that increased significantly 

pie_chart_NIF_m3_inc <- pie_chart(ratio_data = NIF_m3_percent_changes, variable = "Class", column = "Increase", color = "black") +
  labs(
    title = "Percentage of compounds that increased per class level",
  )+
  guides(fill = guide_legend(title = "Compound classes"))
 print(pie_chart_NIF_m3_inc)
 
 pie_chart_NIF_m3_dec <- pie_chart(ratio_data = NIF_m3_percent_changes, variable = "Class", column = "Decrease", color = "black") +
  labs(
    title = "Percentage of compounds that decreased per class level",
  )+
  guides(fill = guide_legend(title = "Compound classes"))
 print(pie_chart_NIF_m3_dec)
 
 
 pie_chart_NIF_m3_all <- pie_chart(ratio_data = NIF_m3_percent_changes, variable = "Class", column = "Significant_Changes", color = "black") +
  labs(
    title = "Percentage of compounds that were significant per class level",
  )+
  guides(fill = guide_legend(title = "Compound classes"))
 print(pie_chart_NIF_m3_all)
 
 
#Intermittent fasting M3 changes as percentages
 #Table 
 IF_m3_percent_changes <- ra_table(fc_data = fold_change_counts_3m_IF, variable = "Class")

# pie chart with percentages of compounds that increased significantly 

pie_chart_IF_m3_inc <- pie_chart(ratio_data = IF_m3_percent_changes, variable = "Class", column = "Increase", color = "black") +
  labs(
    title = "Percentage of compounds that increased per class level",
  )+
  guides(fill = guide_legend(title = "Compound classes"))
 print(pie_chart_IF_m3_inc)
 
 pie_chart_IF_m3_dec <- pie_chart(ratio_data = IF_m3_percent_changes, variable = "Class", column = "Decrease", color = "black") +
  labs(
    title = "Percentage of compounds that decreased per class level",
  )+
  guides(fill = guide_legend(title = "Compound classes"))
 print(pie_chart_IF_m3_dec)
 
 
  pie_chart_IF_m3_all <- pie_chart(ratio_data = IF_m3_percent_changes, variable = "Class", column = "Significant_Changes", color = "black") +
  labs(
    title = "Percentage of compounds that were significant per class level",
  )+
  guides(fill = guide_legend(title = "Compound classes"))
 print(pie_chart_IF_m3_all)
 
 #### fix all of the labels in the legend and try to get percentages to show inside the pie chart. 
 
 
 
```
```{r}
# Comparison between baseline and End of radiation B0 vs follow up 6 and 9

# I have to now change my meta data to make follow up 6 and 9 one single follow up and compare baseline to this "follow up point"

# Not enough replicates to make the comparison




```
# This is for everyone in month 3
```{r}
all_b_m3 <- omu_summary(

  count_data = count_df,
  metadata = meta_data_df,
  numerator = "B0",
  denominator =  "F3",
  response_variable = "Metabolite",
  Factor = "Time",
  log_transform = TRUE,
  p_adjust = "BH",
  test_type = "mwu"
  
)

### See what metabolites are significant 

all_b_m3$diffexpressed <- "NO"
# if Foldchange >1 and pvalue < 0.05, set as "UP"
all_b_m3$diffexpressed[all_b_m3$Fold_Change > 1.0 & all_b_m3$pval < 0.05] <- "UP"
# if Foldchange < -1 and pvalue < 0.05, set as "DOWN"
all_b_m3$diffexpressed[all_b_m3$Fold_Change < 1.0 & all_b_m3$pval < 0.05] <- "DOWN"

#How many significant metabolites? 
all_b_m3 %>% count(diffexpressed)

summary(all_b_m3$Class)## Finding what other classes to subset by

#What are the metabolites that are significant by Class 

all_b_m3_class<- all_b_m3[c("Metabolite", "KEGG","Class", "diffexpressed", "pval","Fold_Change")]
 all_b_m3_class <- all_b_m3_class[which(all_b_m3_class[,"pval"] <= 0.05),]
 
 
### see what class is affected
 summary(all_b_m3_class)
```

# Metaboanalyst 

```{r}
metanr_packages <- function(){
metr_pkgs <- c("impute", "pcaMethods", "globaltest", "GlobalAncova", "Rgraphviz", "preprocessCore", "genefilter", "SSPA", "sva", "limma", "KEGGgraph", "siggenes","BiocParallel", "MSnbase", "multtest", "RBGL", "edgeR", "fgsea", "devtools", "crmn")
list_installed <- installed.packages()
new_pkgs <- subset(metr_pkgs, !(metr_pkgs %in% list_installed[, "Package"]))
if(length(new_pkgs)!=0){if (!requireNamespace("BiocManager", quietly = TRUE))
        install.packages("BiocManager")
        BiocManager::install(new_pkgs)
        print(c(new_pkgs, " packages added..."))
    }

if((length(new_pkgs)<1)){
        print("No new packages added...")
    }
}
```

```{r}
metanr_packages()
```
```{r}
install.packages("https://www.dropbox.com/s/pp9vziji96k5z5k/MetaboAnalystR_3.2.0.tar.gz", repos = NULL, method = "wget")


```


```{r}
install.packages("pacman")

library(pacman)

pacman::p_load(c("impute", "pcaMethods", "globaltest", "GlobalAncova", "Rgraphviz", "preprocessCore", "genefilter", "SSPA", "sva", "limma", "KEGGgraph", "siggenes","BiocParallel", "MSnbase", "multtest", "RBGL", "edgeR", "fgsea"))

```

```{r}
install.packages("devtools")
library(devtools)

# Step 2: Install MetaboAnalystR without documentation
devtools::install_github("xia-lab/MetaboAnalystR", build = TRUE, build_vignettes = FALSE)

```

```{r}
library(MetaboAnalystR)

# Initialize an example data analysis
mSet <- InitDataObjects("conc", "msetora", FALSE)
print("MetaboAnalystR successfully installed and loaded!")

```
#Time Series Analysis 

```{r}
# Want to label the compounds that are defined as other for now using ChemID

count_df <- assign_hierarchy(count_data = count_df, keep_unknowns = TRUE, identifier = "PubChem CID")


```

